---
title: 'Analysis of longitudinal data [DRAFT]'
description: 'Repeated measure data'
---

## Meet and Repeat [DRAFT]

```yaml
type: NormalExercise
key: f9ad1f5b4a
lang: r
xp: 100
skills: 1
```

Welcome to the *Analysis of longitudinal data* chapter.

THIS CHAPTER IS A DRAFT

Exercise 1 introduction: Check your data

`@instructions`
- Read the `bprs` data into memory
- Print out the (column) names of the data
- Look at the structure of the data
- Print out summaries of the variables in the data

`@hint`
- Use `str()` to see structure
- Use `summary()` to compute summaries

`@pre_exercise_code`
```{r}

```

`@sample_code`
```{r}
# read the BPRS data
bprs <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)

# look at the (column) names of human
names(bprs)

# look at the structure of human


# print out summaries of the variables


```

`@solution`
```{r}
# read the human data
bprs <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)

# look at the (column) names of human
names(bprs)

# look at the structure of human
str(bprs)

# print out summaries of the variables
summary(bprs)
```

`@sct`
```{r}

test_function("str", args = "object",incorrect_msg = "Please look at the structure of the bprs data frame")
test_function("summary", args = "object",incorrect_msg = "Please print out summaries of the variables in the BPRS data frame")

test_error()
success_msg("Awsome, you have mastered an important first step of data analysis!")
```

---

## Graphical displays of longitudinal data: The magical gathering

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: e9b52fd18b
```

Exercise 2 introduction: Converting to long form

`@instructions`
- Factor treatment and subject from bprs
- Gather by key = weeks by excluding treatment and subject (how does key work?)
- Mutate a new integer variable week by substringing the only week number
- Glimpse

`@hint`
- Hints

`@pre_exercise_code`
```{r}
library(dplyr); library(tidyr)
bprs <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)
```

`@sample_code`
```{r}
# dplyr & tidyr packages and bprs are available

```

`@solution`
```{r}
# dplyr & tidyr packages and bprs are available

# Factor treatment & subject
bprs$treatment <- factor(bprs$treatment)
bprs$subject <- factor(bprs$subject)

# Convert to long form
bprsl <-  bprs %>% gather(key = weeks, value = bprs, -treatment, -subject)
bprsl <-  bprsl %>% mutate(week = as.integer(substr(weeks,5,5)))

# Let's see the gathered "long" data
glimpse(bprsl)
```

`@sct`
```{r}

success_msg("No tests yet!")
```

---

## Graphical displays of longitudinal data: Individual plots

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: a1be9da8e9
```

Exercise 3 introduction

`@instructions`
- Draw the first plot
- Standardise variable bprs
- Glimpse
- Draw a new plot of the standardised values

`@hint`
- Hints

`@pre_exercise_code`
```{r}
library(dplyr); library(tidyr); library(ggplot2)
bprs <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)
bprs$treatment <- factor(bprs$treatment)
bprs$subject <- factor(bprs$subject)
bprsl <-  bprs %>% gather(key = weeks, value = bprs, -treatment, -subject)
bprsl <-  bprsl %>% mutate(week = as.integer(substr(weeks,5,5)))
rm(bprs)
```

`@sample_code`
```{r}
# dplyr, tidyr & ggplot2 packages and bprsl are available

```

`@solution`
```{r}
# dplyr & tidyr packages and bprs are available

# Plot me! (maybe do this in a function)
p1 <- ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject))
p2 <- p1 + geom_line() + scale_linetype_manual(values = rep(1:10, times=4))
p3 <- p2 + facet_grid(. ~ treatment, labeller = label_both)
p4 <- p3 + theme_bw() + theme(legend.position = "none")
p5 <- p4 + scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
p5

# Standardise

BPRSL <- BPRSL %>%
  group_by(week) %>%
  mutate( stdbprs = (bprs - mean(bprs))/sd(bprs) ) %>%
  ungroup()

# Check what's new
glimpse(BPRSL)

# Plot again
p1 <- ggplot(BPRSL, aes(x = week, y = stdbprs, linetype = subject))
p2 <- p1 + geom_line() + scale_linetype_manual(values = rep(1:10, times=4))
p3 <- p2 + facet_grid(. ~ treatment, labeller = label_both)
p4 <- p3 + theme_bw() + theme(legend.position = "none")
p5 <- p4 + scale_y_continuous(name = "standardized bprs")
p5

```

`@sct`
```{r}

success_msg("No tests yet!")
```
