---
title       : Regression and model validation 
description : Simple regression, multiple regression, regression diagnostics

--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1
## Welcome!

*** =video_link
//player.vimeo.com/video/195829801


--- type:NormalExercise lang:r xp:100 skills:1 key:4fc38a823f
## Formatting data for regression

Welcome to chapter1: Regression and model validation. 

The first step of data analysis with R is reading data into R. Which function and function arguments to use to do this depends on the original format of the data. After the correct function has been identified and data read into R, the data will be in R's `data.frame` format. An $n$ x $d$ data frame has $d$ columns (the variables) and $n$ rows (the observations).

For regression analysis, you choose one variable as the response and some other variable(s) as the predictor(s).


*** =instructions
- instruction

*** =hint
- hint

*** =pre_exercise_code
```{r}

```

*** =sample_code
```{r}


```

*** =solution
```{r}

```

*** =sct
```{r}

test_function("", args = "",incorrect_msg = "")
test_object("", incorrect_msg = "")
test_output_contains("", incorrect_msg = "")

test_error()
success_msg("Good work!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:537efee431
## Simple Linear regression

Exercise info here

*** =instructions
- instruction

*** =hint
- hint

*** =pre_exercise_code
```{r}

```

*** =sample_code
```{r}

```

*** =solution
```{r}

```

*** =sct
```{r}

test_function("", args = "",incorrect_msg = "")
test_object("", incorrect_msg = "")
test_output_contains("", incorrect_msg = "")

test_error()
success_msg("Good work!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:b48fefcaa3
## Multiple linear regression

Exercise info here

*** =instructions
- instruction

*** =hint
- hint

*** =pre_exercise_code
```{r}

```

*** =sample_code
```{r}

```

*** =solution
```{r}

```

*** =sct
```{r}

test_function("", args = "",incorrect_msg = "")
test_object("", incorrect_msg = "")
test_output_contains("", incorrect_msg = "")

test_error()
success_msg("Good work!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:cc303b6569
## Polynomial regression

Exercise info here

*** =instructions
- instruction

*** =hint
- hint

*** =pre_exercise_code
```{r}

```

*** =sample_code
```{r}

```

*** =solution
```{r}

```

*** =sct
```{r}

test_function("", args = "",incorrect_msg = "")
test_object("", incorrect_msg = "")
test_output_contains("", incorrect_msg = "")

test_error()
success_msg("Good work!")
```

