---
title: '[UNDER CONSTRUCTION] Analysis of longitudinal data'
description: 'Graphical Displays and Summary Measure Approach, Linear Mixed Effects Models for Normal Response Variables'
---

## [UNDER CONSTRUCTION] Meet and Repeat: PART I

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: 7f7f6ab8ba
```

Welcome to the *Analysis of longitudinal data* chapter.

Many studies in the behavioral sciences involve several measurement or observations of the response variable of interest on each subject in the study. For example, the response variable may be measured under a number of different experimental conditions or on a number of different occasions over time; such data are labelled repeated measures or *longitudinal data*. In the first part (I) of this chapter useful methods for the graphical exploration of this type of data are described and a simple method for their analysis are introduced, with the warning that although simple the method should be used only in the initial stage of dealing with the data; more appropriate methods will be discussed in part II.

In the first part we will dwelve in to the BPRS data [SITATION], in which 40 male subjects were randomly assigned to one of two treatment groups and each subject was rated on the brief psychiatric rating scale (BPRS) measured before treatment began (week 0) and then at weekly intervals for eight weeks. The BPRS assesses the level of 18 symptom constructs such as hostility, suspiciousness, hallucinations and grandiosity; each of these is rated from one (not present) to seven (extremely severe). The scale is used to evaluate patients suspected of having schizophrenia.

`@instructions`
- Read the `BPRS` data into memory
- Print out the (column) names of the data
- Look at the structure of the data
- Print out summaries of the variables in the data
- Pay special attention to the structure of the data

`@hint`
- Use `str()` to see structure
- Use `summary()` to compute summaries

`@pre_exercise_code`
```{r}

```

`@sample_code`
```{r}
# Read the BPRS data
BPRS <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)

# Look at the (column) names of BPRS
names(BPRS)

# Look at the structure of BPRS


# Print out summaries of the variables


```

`@solution`
```{r}
# Read the BPRS data
BPRS <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)

# Look at the (column) names of BPRS
names(BPRS)

# Look at the structure of BPRS
str(BPRS)

# print out summaries of the variables
summary(BPRS)

```

`@sct`
```{r}
test_function("str", args = "object",incorrect_msg = "Please look at the structure of the BPRS data frame")
test_function("summary", args = "object",incorrect_msg = "Please print out summaries of the variables in the BPRS data frame")
test_error()
success_msg("Awsome, you have mastered an important first step of data analysis!")
```

---

## Graphical displays of longitudinal data: The magical gather()

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: 4751cc12a0
```

To be able to study the possible diffecences in the bprs value between the treatment groups and the possible change of the value in time, we don't want the weeks to be individual variables. In comes the gather() function which you can probably remember from previous chapters. The gather() function takes multiple columns and collapses them into key-value pairs, so that we can have the weeks as values of a new variable week. You can find more information about gather in the package documentation with ?gather or in the dplyr cheatsheet.

`@instructions`
- Factor variables treatment and subject
- Use `gather()` to convert BPRS to a long form
- Use `mutate()` and `substring()` to create a new integer variable `week`
- Glimpse the data using `glimpse()`

`@hint`
- Hints

`@pre_exercise_code`
```{r}
BPRS <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)
```

`@sample_code`
```{r}
# The data BPRS is available

# Access the packages dplyr and tidyr
library(dplyr)
library(tidyr)

# Factor treatment & subject
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)

# Convert to long form
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
BPRSL <-  BPRSL %>% mutate(week = as.integer(substr(weeks,5,5)))

# Take a glimpse at the BPRSL data
glimpse(BPRSL)

```

`@solution`
```{r}
# The data BPRS is available

# Access the packages dplyr and tidyr
library(dplyr)
library(tidyr)

# Factor treatment & subject
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)

# Convert to long form
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
BPRSL <-  BPRSL %>% mutate(week = as.integer(substr(weeks,5,5)))

# Take a glimpse at the BPRSL data
glimpse(BPRSL)

```

`@sct`
```{r}
success_msg("No tests")
```

---

## Graphical displays of longitudinal data: Individual plots

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: 5d54d96aee
```

Graphical displays of data are almost always useful for exposing patterns in the data, particularly when these are unexpected; this might be of great help in suggesting which class of models might be most sensibly applied in the later more formal analysis.

To begin we shall plot the BPRS values for all 40 men, differentiating between the treatment groups into which the men have been randomized. This simple graph makes a number of features of the data readily apparent.

REMEMBER: In `ggplot2` or `dplyr` syntax, you generally do not need to "quote" variable names!

`@instructions`
- Draw the plot with `week` on the x-axis and `bprs` on the y-axis
- Inspect the plot. See how both the BPRS-score and the variability between individuals decrease over the eight weeks time

`@hint`
- Hints

`@pre_exercise_code`
```{r}
library(dplyr); library(tidyr)
BPRS <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
BPRSL <-  BPRSL %>% mutate(week = as.integer(substr(weeks,5,5)))
rm(BPRS)
```

`@sample_code`
```{r}
# dplyr, tidyr packages and BPRSL are available

#Access the package ggplot2
library(ggplot2)

# Draw the plot
ggplot(BPRSL, aes(x = "Change me!", y = "Change me too!", linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```

`@solution`
```{r}
# dplyr, tidyr packages and BPRSL are available

#Access the package ggplot2
library(ggplot2)

# Draw the plot
ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```

`@sct`
```{r}
success_msg("Success!")
```

---

## Graphical displays of longitudinal data: Standardise

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: 2923bbb346
```

An important effect we want to take notice is how the men who have higher BPRS values at the beginning tend to have higher values throughout the study. This phenomenon is generally referred to as tracking.

The tracking phenomenon can be seen more clearly in a plot of the standardized values of each
observation, i.e., the values obtained by subtracting the relevant occasion mean from the original observation and then dividing by the corresponding visit standard deviation.

$$standardised(x) = \frac{x - mean(x)}{ sd(x)}$$

REMEMBER: In `ggplot2` or `dplyr` syntax, you generally do not need to "quote" variable names!

`@instructions`
- Assign `week` as the grouping variable
- Standardise the variable `bprs`
- Glimpse the data now with the standardised `brps`
- Plot the data now with the standardised `brps`

`@hint`
- Hints

`@pre_exercise_code`
```{r}
library(dplyr); library(tidyr); library(ggplot2)
BPRS <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
BPRSL <-  BPRSL %>% mutate(week = as.integer(substr(weeks,5,5)))
rm(BPRS)
```

`@sample_code`
```{r}
# dplyr, tidyr and ggplot2 packages and BPRSL are available

# Standardise the variable bprs
BPRSL <- BPRSL %>%
  group_by("Change me!") %>%
  mutate(stdbprs = "Change me!") %>%
  ungroup()

# Glimpse the data
glimpse(BPRSL)

# Plot again with the standardised bprs
ggplot(BPRSL, aes(x = week, y = stdbprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  scale_y_continuous(name = "standardized bprs")

```

`@solution`
```{r}
# dplyr, tidyr and ggplot2 packages and BPRSL are available

# Standardise the variable bprs
BPRSL <- BPRSL %>%
  group_by(week) %>%
  mutate(stdbprs = (bprs - mean(bprs))/sd(bprs) ) %>%
  ungroup()

# Glimpse the data
glimpse(BPRSL)

# Plot again with the standardised bprs
ggplot(BPRSL, aes(x = week, y = stdbprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  scale_y_continuous(name = "standardized bprs")

```

`@sct`
```{r}
success_msg("Success!")
```

---

## Graphical displays of longitudinal data: Summary graphs

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: c2cb0fb62a
```

With large numbers of observations, graphical displays of individual response profiles are of little use and investigators then commonly produce graphs showing average (mean) profiles for each treatment group along with some indication of the variation of the observations at each time point, in this case the standard error of mean

$$se = \frac{sd(x)}{\sqrt{n}}$$

`@instructions`
- Create the summary data `BPRSS` with the mean and standard error of the variable `bprs`
- Glimpse the data
- Plot the mean profiles (with `geom_errorbar()` line commented out)
- Uncomment the `geom_errorbar()` line and plot the mean profiles again
- Note the considerable overlap in the mean profiles of the two
treatment groups suggesting there might be little difference between
the two groups in respect to the mean BPRS values

`@hint`
- Hints

`@pre_exercise_code`
```{r}
library(dplyr); library(tidyr); library(ggplot2)
BPRS <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
BPRSL <-  BPRSL %>% mutate(week = as.integer(substr(weeks,5,5)))
rm(BPRS)
BPRSL <- BPRSL %>%
  group_by(week) %>%
  mutate( stdbprs = (bprs - mean(bprs))/sd(bprs) ) %>%
  ungroup()
```

`@sample_code`
```{r}
# dplyr, tidyr & ggplot2 packages and BPRSL are available

# Number of weeks, baseline (week 0) included
n <- BPRSL$week %>% unique() %>% length()

# Summary data with mean and standard error of bprs by treatment and week 
BPRSS <- BPRSL %>%
  group_by(treatment, week) %>%
  summarise( mean = "Change me!", se = "Change me!" ) %>%
  ungroup()

# Glimpse the data
glimpse(BPRSS)

# Plot the mean profiles
ggplot(BPRSS, aes(x = week, y = mean, linetype = treatment, shape = treatment)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2)) +
  #geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.8)) +
  scale_y_continuous(name = "mean(bprs) +/- se(bprs)")

```

`@solution`
```{r}
# dplyr, tidyr & ggplot2 packages and BPRSL are available

# Number of weeks, baseline (week 0) included
n <- BPRSL$week %>% unique() %>% length()

# Summary data with mean and standard error of bprs by treatment and week 
BPRSS <- BPRSL %>%
  group_by(treatment, week) %>%
  summarise( mean = mean(bprs), se = sd(bprs)/sqrt(n) ) %>%
  ungroup()

# Glimpse the data
glimpse(BPRSS)

# Plot the mean profiles
ggplot(BPRSS, aes(x = week, y = mean, linetype = treatment, shape = treatment)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.8)) +
  scale_y_continuous(name = "mean(bprs) +/- se(bprs)")

```

`@sct`
```{r}
success_msg("Success!")
```

---

## Graphical displays of longitudinal data: Find the outlaw

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: 539b15366d
```

As an example of the summary measure approach we will look into the post treatment values of the BPRS. The mean of weeks 1 to 8 will be our summary measure. First calculate this measure and then look at boxplots of the measure for each treatment group. See how the mean summary measure is more variable in the second treatment group and its distribution in this group is somewhat skew. The boxplot of the second group also reveals an outlier, a subject whose mean BPRS score of the eight weeks is over 70. It might bias the conclusions from further comparisons of the groups, so we shall remove that subject from the data. Without the outlier, try to figure which treatment group might have the lower the eight-week mean. Think, considering the variation, how can we be sure? 


`@instructions`
- Create the summary data BPRSL8S
- Glimpse the data
- Draw the boxplot and observe the outlier
- Find a suitable threshold value and use `filter()` to exclude the outlier to form a new data BPRSL8S1
- Glimpse and draw a boxplot of the new data to check the outlier has been dealt with

`@hint`
- Hints

`@pre_exercise_code`
```{r}
library(dplyr); library(tidyr); library(ggplot2)
BPRS <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
BPRSL <-  BPRSL %>% mutate(week = as.integer(substr(weeks,5,5)))
rm(BPRS)
BPRSL <- BPRSL %>%
  group_by(week) %>%
  mutate( stdbprs = (bprs - mean(bprs))/sd(bprs) ) %>%
  ungroup()
```

`@sample_code`
```{r}
# dplyr, tidyr & ggplot2 packages and BPRSL are available

# Create a summary data by treatment and subject with mean as the summary variable (ignoring baseline week 0).
BPRSL8S <- BPRSL %>%
  filter(week > 0) %>%
  group_by(treatment, subject) %>%
  summarise( mean=mean(bprs) ) %>%
  ungroup()

# Glimpse the data
glimpse(BPRSL8S)

# Draw a boxplot of the mean versus treatment
ggplot(BPRSL8S, aes(x = treatment, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(bprs), weeks 1-8")

# Create a new data by filtering the outlier
BPRSL8S1 <- "Change me!"

```

`@solution`
```{r}
# dplyr, tidyr & ggplot2 packages and BPRSL are available

# Create a summary data by treatment and subject with mean as the summary variable (ignoring baseline week 0).
BPRSL8S <- BPRSL %>%
  filter(week > 0) %>%
  group_by(treatment, subject) %>%
  summarise( mean=mean(bprs) ) %>%
  ungroup()

# Glimpse the data
glimpse(BPRSL8S)

# Draw a boxplot of the mean versus treatment
ggplot(BPRSL8S, aes(x = treatment, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(bprs), weeks 1-8")

# Create a new data by filtering the outlier
BPRSL8S1 <- BPRSL8S %>%
  filter(mean < 60)

```

`@sct`
```{r}
success_msg("Success!")
```

---

## Graphical displays of longitudinal data: T-test and anova

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: '4714162745'
```

Although the informal graphical material presented up to now has all indicated a lack of difference in the two treatment groups, most investigators would still require a formal test for a difference. Consequently we shall now apply a t-test to assess any difference between the treatment groups, and also calculate a confidence interval for this difference. We use the data without the outlier created in the previous exercise. The t-test confirms the lack of any evidence for a group difference. Also the 95% confidence interval is wide and includes the zero, allowing for similar conclusions to be made.

Baseline measurements of the outcome variable in a longitudinal study are often correlated with the chosen summary measure and using such measures in the analysis can often lead to substantial gains in precision when used appropriately as a covariate in an analysis of covariance. We can illustrate the analysis on the data using the BPRS value corresponding to time zero taken prior to the start of treatment as the baseline covariate. We see that the baseline BPRS is strongly related to the BPRS values taken after treatment has begun, but there is still no evidence of a treatment difference even after conditioning on the baseline value.

`@instructions`
- Perform a two-sample t-test...
- Add the baseline from the original data as a new variable to the summary data
- Fit the linear model...
- Compute the analysis of variance table for the fitted model

`@hint`
- Hints

`@pre_exercise_code`
```{r}
library(dplyr); library(tidyr); library(ggplot2)
BPRS <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
BPRSL <-  BPRSL %>% mutate(week = as.integer(substr(weeks,5,5)))
BPRSL <- BPRSL %>%
  group_by(week) %>%
  mutate( stdbprs = (bprs - mean(bprs))/sd(bprs) ) %>%
  ungroup()
BPRSL8S <- BPRSL %>%
  filter(week > 0) %>%
  group_by(treatment, subject) %>%
  summarise( mean=mean(bprs) ) %>%
  ungroup()
rm(BPRSL)
BPRSL8S1 <- BPRSL8S %>%
  filter(mean < 60)
```

`@sample_code`
```{r}
# dplyr, tidyr & ggplot2 packages and BPRSL8S & BPRSL8S1 data are available

# Perform a two-sample t-test
t.test(mean ~ treatment, data = BPRSL8S1, var.equal = TRUE)

# Add the baseline from the original data as a new variable to the summary data
BPRSL8S2 <- BPRSL8S %>%
  mutate(baseline = BPRS$week0)

# Fit the linear model with the mean as the response 
fit <- lm(mean ~ baseline + treatment, data = BPRSL8S2)

# Compute the analysis of variance table for the fitted model
anova(fit)

```

`@solution`
```{r}
# dplyr, tidyr & ggplot2 packages and BPRSL8S & BPRSL8S1 data are available

# Perform a two-sample t-test
t.test(mean ~ treatment, data = BPRSL8S1, var.equal = TRUE)

# Add the baseline from the original data as a new variable to the summary data
BPRSL8S2 <- BPRSL8S %>%
  mutate(baseline = BPRS$week0)

# Fit the linear model with the mean as the response 
fit <- lm(mean ~ baseline + treatment, data = BPRSL8S2)

# Compute the analysis of variance table for the fitted model
anova(fit)

```

`@sct`
```{r}
success_msg("Success!")
```

---

## Meet and Repeat: PART II

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: 707255f740
```

Welcome to the PART II of *Analysis of longitudinal data* chapter.

Longitudinal data, where a response variable is measured on each subject on several different occasions poses problems for their analysis because the repeated measurements on each subject are very likely to be correlated rather than independent. In PART II of this chapter methods for dealing with longitudinal data which aim to account for the correlated nature of the data and where the response is assumed to be normally distributed are discussed.

To investigate the use of linear mixed effects models in practice, we shall use data from a nutrition study conducted in three groups of rats. The groups were put on different diets, and each animal’s body weight (grams) was recorded repeatedly (approximately) weekly, except in week seven when two recordings were taken) over a 9-week period. The question of most interest is whether the growth profiles of the three groups differ.


`@instructions`
- Read the `RATS` data into memory
- Glimpse the data

`@hint`
- Hints

`@pre_exercise_code`
```{r}
library(dplyr)
```

`@sample_code`
```{r}
# dplyr is available

# read the RATS data
RATS <- read.table("http://assets.datacamp.com/production/repositories/708/datasets/9475c75ae0ffa340f53ef895e2494b256774a1d6/rats.txt", header = TRUE, sep = '\t')

# Factor variables ID and Group
RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)

# Glimpse the data
glimpse(RATS)

```

`@solution`
```{r}
# dplyr is available

# read the RATS data
RATS <- read.table("http://assets.datacamp.com/production/repositories/708/datasets/9475c75ae0ffa340f53ef895e2494b256774a1d6/rats.txt", header = TRUE, sep = '\t')

# Factor variables ID and Group
RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)

# Glimpse the data
glimpse(RATS)

```

`@sct`
```{r}
success_msg("Success!")
```

---

## Linear Mixed Effects Models: Gather 'round

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: f2f670e284
```

Again, to be able to study the diffecences between the variable of interest, that is the weight of the individual rats, and the groups as well as the change of the weight in time, we want to *gather* the data to a long form.

`@instructions`
- Assign `key` as `WD` and value as `Weight` and convert the data to a long form
- `glimpse` the data

`@hint`
- Hint

`@pre_exercise_code`
```{r}
library(dplyr); library(tidyr)
RATS <- read.table("http://assets.datacamp.com/production/repositories/708/datasets/9475c75ae0ffa340f53ef895e2494b256774a1d6/rats.txt", header = TRUE, sep = '\t')
RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)
```

`@sample_code`
```{r}
# dplyr, tidyr and RATS are available

# Convert data to long form
RATSL <- RATS %>% gather(key = "Change me!", value = "Change me!", -ID, -Group) %>%
  mutate(Time = as.integer(substr(WD,3,4))) 

# Glimpse the data
glimpse(RATSL)

```

`@solution`
```{r}
# dplyr, tidyr and RATS are available

# Convert data to long form
RATSL <- RATS %>% gather(key = WD, value = Weight, -ID, -Group) %>%
  mutate(Time = as.integer(substr(WD,3,4))) 

# Glimpse the data
glimpse(RATSL)

```

`@sct`
```{r}
success_msg("Success!")
```

---

## Linear Mixed Effects Models: Plot first, ask questions later

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: e85d52657c
```

To begin, we shall ignore the repeated-measures structure of the data and assume that all the observations are independent of one another. Now if we simply ignore that the sets of 11 weights come from the same rat, we have a data set consisting of 176 weights, times, and group memberships that we see can easily be analyzed using multiple linear regression. To begin, we will plot the data, identifying the observations in each group but ignoring the longitudinal nature of the data.

`@instructions`
- Check the dimensions of RATSL
- Observe the difference between the weights of the groups 1 rats and those in the other two groups

`@hint`
- Hint

`@pre_exercise_code`
```{r}
library(dplyr); library(tidyr); library(ggplot2)
RATS <- read.table("http://assets.datacamp.com/production/repositories/708/datasets/9475c75ae0ffa340f53ef895e2494b256774a1d6/rats.txt", header = TRUE, sep = '\t')
RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)
RATSL <- gather(RATS, key = WD, value = Weight, -ID, -Group) %>%
  mutate(Time = as.integer(substr(WD,3,4))) 
```

`@sample_code`
```{r}
# dplyr, tidyr and RATSL are available

# Check the dimensions of the data

# Plot the RATSL data
ggplot(RATSL, aes(x = Time, y = Weight, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Weight (grams)") +
  theme(legend.position = "top")
```

`@solution`
```{r}
# dplyr, tidyr and RATSL are available

# Check the dimensions of the data
dim(RATSL)

# Plot the RATSL data
ggplot(RATSL, aes(x = Time, y = Weight, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Weight (grams)") +
  theme(legend.position = "top")
```

`@sct`
```{r}
success_msg("Success!")
```

---

## Linear Mixed Effects Models: Linear Model (assumed independence)

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: 3c0a95b7b1
```

Exercise 10 introduction

`@instructions`
- Instructions

`@hint`
- Hint

`@pre_exercise_code`
```{r}
library(dplyr); library(tidyr)
RATS <- read.table("http://assets.datacamp.com/production/repositories/708/datasets/9475c75ae0ffa340f53ef895e2494b256774a1d6/rats.txt", header = TRUE, sep = '\t')
RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)
RATSL <- gather(RATS, key = WD, value = Weight, -ID, -Group) %>%
  mutate(Time = as.integer(substr(WD,3,4))) 
```

`@sample_code`
```{r}
# dplyr, tidyr RATS and RATSL are available

# sample code here

```

`@solution`
```{r}
# dplyr, tidyr, RATS and RATSL are available

# dummies (in Table) vs summary output: D1 = Group2, D2 = Group3
RATS_reg <- lm(Weight ~ Time + Group, data = RATSL)
summary(RATS_reg)

```

`@sct`
```{r}
success_msg("Success!")
```

---

## Linear Mixed Effects Models: Random Intercept Model

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: 98fbbfad2e
```

Exercise 11 introduction

`@instructions`
- Instructions

`@hint`
- Hint

`@pre_exercise_code`
```{r}
library(dplyr); library(tidyr); library(lme4)
RATS <- read.table("http://assets.datacamp.com/production/repositories/708/datasets/9475c75ae0ffa340f53ef895e2494b256774a1d6/rats.txt", header = TRUE, sep = '\t')
RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)
RATSL <- gather(RATS, key = WD, value = Weight, -ID, -Group) %>%
  mutate(Time = as.integer(substr(WD,3,4))) 
```

`@sample_code`
```{r}
# dplyr, tidyr, lme4, RATS and RATSL are ava

# sample code here

```

`@solution`
```{r}
# dplyr, tidyr, lme4, RATS and RATSL are available

# Table 9.4

RATS_ref <- lmer(Weight ~ Time + Group + (1 | ID), data = RATSL, REML = FALSE)
summary(RATS_ref)

# Table 9.5

RATS_ref1 <- lmer(Weight ~ Time + Group + (Time | ID), data = RATSL, REML = FALSE)
summary(RATS_ref1)

anova(RATS_ref1, RATS_ref)

# Table 9.6

RATS_ref2 <- lmer(Weight ~ Time * Group + (Time | ID), data = RATSL, REML = FALSE)
summary(RATS_ref2)

anova(RATS_ref1, RATS_ref2)


```

`@sct`
```{r}
success_msg("Success!")
```
