---
  title       : Dimensionality reduction techniques 
description : Principal component analysis (PCA), Correspondence analysis (CA)

--- type:NormalExercise lang:r xp:100 skills:1 key:00f0a592f0
## Meet the human data

The human dataset is a combination of the indicators behind the human development index and the gender inequality index. The data originates from the United Nations Development Programme. See [their data page](http://hdr.undp.org/en/content/human-development-index-hdi) for more information. For a nice overview and detailed explanations see the [calculating the human development indices pdf](http://hdr.undp.org/sites/default/files/hdr2015_technical_notes.pdf).

Some of the variable names have been shortened but the connections should be clear. "F" always stands for female and "M" for male. There are two new variables:

- **Edu2.FM**: The ratio of the percentages of female and male populations with secondary education (Edu2.F / Edu2.M)
- **Labo.FM**: The ratio of the percentages of female and male populations in the labour force. (Labo.F / Labo.M)

*** =instructions
- Read the human data into memory
- Print out the (column) names of the data
- Look at the structure of the data
- Print out summaries of the variables in the data

*** =hint
- Use `str()` to see structure
- Use `summary()` to compute summaries

*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
# read the human data
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human1.txt", sep  =",", header = T)

# look at the (column) names of human
names(human)

# look at the structure of human


# print out summaries of the variables


```

*** =solution
```{r}
# read the human data
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human1.txt", sep  =",", header = T)

# look at the (column) names of human
names(human)

# look at the structure of human
str(human)

# print out summaries of the variables
summary(human)
```

*** =sct
```{r}

test_function("str", args = "object",incorrect_msg = "Please look at the structure of the human data frame")
test_function("summary", args = "object",incorrect_msg = "Please print out summaries of the variables in the human data frame")

test_error()
success_msg("Awsome, you have mastered an important first step!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:63c85a3cfc
## String manipulation

Sometimes a variable is coded in a way that is not natural for R to understand. For example, large integers can sometimes be coded with a comma to separate thousands (1,000 == 1000). In these cases, R interprets the variable as a **factor** or a **character.** 

In some cases you could use the `dec` argument in `read.table()` to get around this, but if the data also includes decimals separated by a dot, this is not an option. To get rid of the unwanted commas, we need *string manipulation*.

In programming, a combination of characters is usually called a string. In R, strings are of the basic type character and they can be created by using quotation marks or specific functions.  

There are quite a few functions in Base R that can be used to manipulate characters, but there is also a bit more consintent and simple tidyverse package **stringr.**


*** =instructions
- Access the stringr package
- Print out a summary of the Gross National Income (GNI) variable in `human`
- Execute the sample code where the comma is removed from each value of GNI.
- Adjust the code: Use the pipe operator (`%>%`) to convert the resulting vector to numeric with `as.numeric`.

*** =hint
- Use `$` to access a single column of a data frame.
- Use `str()` to look at the structure of any object
- Add the pipe and `as.numeric` to the row where `str_replace()` is used

*** =pre_exercise_code
```{r}
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human1.txt", sep  =",", header = T)
```

*** =sample_code
```{r}
# human is available

# access the stringr package
library(stringr)

# look at the structure of the GNI column in 'human'


# remove the commas from GNI and convert it to numeric
str_replace(human$GNI, pattern=",", replace ="")


```

*** =solution
```{r}
# human is available

# access the stringr package
library(stringr)

# look at the structure of the GNI variable in human
str(human$GNI)

# remove the commas from GNI and convert to numeric
str_replace(human$GNI, pattern=",", replace ="") %>% as.numeric


```

*** =sct
```{r}

test_function("str", args = "object",incorrect_msg = "Please look at the structure of GNI in human")
test_output_contains("str_replace(human$GNI, pattern=',', replace ='') %>% as.numeric", incorrect_msg = "Please adjust the code to convert the gni variable to numeric and print out the result")
 
test_error()
success_msg("Wow, what a manipulator you are! You certainly can have your way with words now.")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:99d98acadf
## Dealing with not available (NA) values

In R, NA stands for not available, which means that the data point is missing. If a variable you wish to analyse contains missing values, there are usually two main options:

- Remove the observations with missing values
- Replace the missing values with actual values using an *imputation* technique.

We will use the first option, which is the simplest solution.

*** =instructions
- Create a smaller version of the human data by selecting the variables defined in `keep`
- Use complete.cases() on human to print out a logical "completeness indicator" vector
- Print out the 'human' data along with an added completeness indicator. When is the indicator FALSE and when is it TRUE? You can also look at the help page of `complete.cases()`.
- `filter()` out all the rows with any `NA` values. Right now, `TRUE` is simply recycled so that nothing is filtered out.

*** =hint
- Use `complete.cases()` on 'human' again to define the 'complete' column
- Use the logical vector created by complete.cases to filter out the rows with `NA` values.

*** =pre_exercise_code
```{r}
library(dplyr)
# read data
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human1.txt", sep  =",", header = T)
human$GNI <- gsub(",", "", human$GNI) %>% as.numeric
```

*** =sample_code
```{r}
# human with modified GNI and dplyr are available

# columns to keep
keep <- c("Country", "Edu2.FM", "Labo.FM", "Life.Exp", "Edu.Exp", "GNI", "Mat.Mor", "Ado.Birth", "Parli.F")

# select the 'keep' columns
human <- select(human, one_of(keep))

# print out a completeness indicator of the 'human' data
complete.cases(human)

# print out the data along with a completeness indicator
data.frame(human[-1], comp = "change me!")

# filter out all rows with NA values
human_ <- filter(human, TRUE)

```

*** =solution
```{r}
# human with modified GNI and dplyr are available

# columns to keep
keep <- c("Country", "Edu2.FM", "Labo.FM", "Life.Exp", "Edu.Exp", "GNI", "Mat.Mor", "Ado.Birth", "Parli.F")

# select the 'keep' columns
human <- select(human, one_of(keep))

# print out a completeness indicator of the 'human' data
complete.cases(human)

# print out the data along with a completeness indicator as the last column
data.frame(human[-1] , comp = complete.cases(human))

# filter out all rows with NA values
human_ <- filter(human, complete.cases(human))

```

*** =sct
```{r}

test_output_contains("complete.cases(human)")
test_output_contains("cbind(human , complete = complete.cases(human))", 
                     incorrect_msg = "Please print out the human data along with the completeness indicators")
test_object("human_", incorrect_msg = "Please filter out the incomplete observations from the human data and create object `human_`")

test_error()
success_msg("Great work! You complete me <3")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:17ccfcda8e
## Excluding observations

Besides missing values, there might be other reasons to exclude observations. In our human data, there are a few data points which have been computed from other observations. We want to remove them before further analysis.

The basic way in R to reference the rows or columns of a data frame is to use brackets (`[,]`) along with indices or names. When brackets are used with data frames, a comma is used to separate row and column references. In the examples below, `df` is a data frame.

```
df[,] # select every row and every column
df[1:5, ] # select first five rows
df[, c(2, 5)] # select 2nd and 5th columns
```

*** =instructions
- Use `tail()` to print out the last 10 observations of human (hint: `?tail`)
- Create object `i_last`
- Create data frame 'human_' by selecting rows from the 1st to `i_last` from `human`.
- Define the rownames in 'human_' by the Country column

*** =hint
- Use `1:i_last` to select rows from 1 to `i_last`

*** =pre_exercise_code
```{r}
library(dplyr)
# read data
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human1.txt", sep  =",", header = T)
human$GNI <- gsub(",", "", human$GNI) %>% as.numeric
keep <- c("Country", "Edu2.FM", "Labo.FM", "Life.Exp", "Edu.Exp", "GNI", "Mat.Mor", "Ado.Birth", "Parli.F")
human <- select(human, one_of(keep))
human <- filter(human, complete.cases(human))

```

*** =sample_code
```{r}
# human without NA is available


# look at the last 10 observations


# define the last indice we want to keep
i_last <- nrow(human) - 7

# choose everything until the last 7 observations
human_ <- human["change me!", ]

# add countries as rownames
rownames(human_) <- human_$Country

```

*** =solution
```{r}
# human without NA is available

# add countries as rownames
rownames(human) <- human$Country

# look at the last 10 observations
tail(human, 10)

# last indice we want to keep
i_last <- nrow(human) - 7

# choose everything until the last 7 observations
human_ <- human[1:i_last, ]

# add countries as rownames
rownames(human) <- human$Country

```

*** =sct
```{r}

test_function("tail", args = c("x", "n"))
test_object("human_", incorrect_msg = "Please create the data frame 'human_' as instructed by selecting rows from 'human'")

test_error()
success_msg("Yes! You're definately included in awsomeland!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:b94de4cad1
## Exploring the human countries

A pairs plot or a generalized pairs plot is a good way of visualizing a reasonably sized data frame. Correlations can be computed with the `cor()` function and visualized with the 

*** =instructions
- Create the data frame 'human_' by removing the Country variable from 'human'
- Draw a scatter plot matrix of the 'human_' variables with `pairs()`. Make the points more colorfull by adding the argument `col = "blue"`.
- Compute and print out the correlation matrix of 'human_'
- Adjust the code: use the pipe operator (`%>%`) and visualize the correlation matrix with `corrplot()`.

*** =hint
- The pipe assigns the output on its left as the first argument to the function name on its right. Use it on the same or where the correlation matrix is computed


*** =pre_exercise_code
```{r}
library(dplyr)
# read data
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human1.txt", sep  =",", header = T)
human$GNI <- gsub(",", "", human$GNI) %>% as.numeric
keep <- c("Country", "Edu2.FM", "Labo.FM", "Life.Exp", "Edu.Exp", "GNI", "Mat.Mor", "Ado.Birth", "Parli.F")
human <- select(human, one_of(keep))
human <- filter(human, complete.cases(human))
rownames(human) <- human$Country
n_last <- nrow(human) - 7
human <- human[1:n_last, ]

# corrplot
source("https://raw.githubusercontent.com/taiyun/corrplot/master/R/corrplot.R")
source("https://raw.githubusercontent.com/taiyun/corrplot/master/R/colorlegend.R")
source("https://raw.githubusercontent.com/taiyun/corrplot/master/R/corrMatOrder.R")
source("https://raw.githubusercontent.com/taiyun/corrplot/master/R/corrRect.R")
source("https://raw.githubusercontent.com/taiyun/corrplot/master/R/corrRect.hclust.R")
source("https://raw.githubusercontent.com/taiyun/corrplot/master/R/corrplot.mixed.R")

```

*** =sample_code
```{r}
# modified human, dplyr and the corrplot functions are available

# remove the Country variable
human_ <- select(human, -Country)

# draw a scatter plot matrix of human_ with pairs()


# compute the correlation matrix and visualize it with corrplot
cor(human_)



```

*** =solution
```{r}
# modified human, dplyr and the corrplot functions are available

# remove the Country variable
human_ <- select(human, -Country)

# draw a scatter plot matrix with pairs()
pairs(human_, col = "blue")

# compute the correlation matrix and visualize it with corrplot
cor(human_) %>% corrplot


```

*** =sct
```{r}

test_function("pairs", args = "x", incorrect_msg = "Please use the `pairs()` function on `human_`")
test_pipe(absent_msg = "Please use the pipe operator to apply the `corrplot()` function on the correlation matrix of 'human_'")

test_error()
success_msg("Great work!!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:42c18276a9
## PCA with R

There are two functions in the default package distribution of R that can be used to perform Principal Components Analysis (PCA): `princomp()` and `prcomp()`. This is because PCA can be performed by two sightly different matrix decomposition methods from linear algebra: the Eigenvalue Decomposition and the Singular Value Decomposition (SVD). 

The `prcomp()` function uses the SVD and is the preferred and more numerically accurate method for PCA.

Both methods quite literally *decompose* a data matrix into a product of smaller matrices, which let's us extract the underlying *principal components*. This makes it possible to approximate a lower dimensional representation of the data by choosing only a few principal components.


*** =instructions
- Create 'human_std' by standardizing the variables in 'human'.
- Print out summaries of the standardized variables. What are the means? Do you know the standard deviations?
- Use `prcomp()` to perform principal component analysis on the standardized data. Save the results in the object `pca_human`
- Use `biplot()` to draw a biplot of `pca_human`.
- Experiment with the argument `cex` which should be a vector of length 2 and can be used to scale the labels in the biplot. Which number affects what? Try for example `cex = c(0.5, 0.8)`. 
- Add the argument `col = c("grey40", "deeppink2")`


*** =hint
- Use the `summary() function to compute summeries of the variables in a data frame

*** =pre_exercise_code
```{r}
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep  =",", header = T)

```

*** =sample_code
```{r}
# modified human is available

# standardize the variables
human_std <- scale(human)

# print out summaries of the standardized variables


# perform principal component analysis (with the SVD method)
pca_human <- prcomp(human_std)


# draw a biplot of the principal components and the original variables
biplot(pca_human)

```

*** =solution
```{r}
# modified human is available

# standardize the variables
human_std <- scale(human)

# print out summaries of the standardized variables
summary(human_std)

# perform principal component analysis (with the SVD method)
pca_human <- prcomp(human_std)

# draw a biplot of the 2-dimensional principal component representation of the observations
biplot(pca_human, cex = c(0.5, 0.8), col = c("grey40", "deeppink2"))

```

*** =sct
```{r}

test_function("summary", args = "object")

test_error()
success_msg("Great work!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:65a759a329
## A biplot of PCA

Exercise info here

*** =instructions
- Create and print out a summary of `pca_human` (created in the previous exercise)
- Create object `pca_pr` and print it out
- Adjust the code: instead of proportions, save the percentages of variance in the `pca_pr` object. Round the percentages to 1 digit.
- Execute the `paste0()` function and then save the output to a new object `pc_lab`.
- Draw the biplot again. Use the first value of the `pc_lab` vector as the label for the x-axis and the second value as the label for the y-axis.

*** =hint
- Percentages are proportions on a different scale. Multiplication by 100 changes proportions to percentages.
- Objects are created with the assign operator `<-`
- Brackets can be used to access values of a vector: `V[1]` gets the first value of `V`

*** =pre_exercise_code
```{r}
library(dplyr)
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep  =",", header = T)
human_std <- scale(human)
pca_human <- prcomp(human_std)
biplot(pca_human, cex = c(0.5, 0.8), col = c("grey40", "deeppink2"))

```

*** =sample_code
```{r}
# pca_human, dplyr are available

# create a summary of pca_human
s <- summary(pca_human)

# print out the summary


# rounded percetanges of variance captured by each PC
pca_pr <- round(1*s$importance[2, ], digits = 5)

# print out the percentages of variance


# create object pc_lab to be used as axis labels
paste0("PC", 1:8, " (", pca_pr, "%)")

# draw a biplot
biplot(pca_human, cex = c(0.5, 0.8), col = c("grey40", "deeppink2"), xlab = NA, ylab = NA)

```

*** =solution
```{r}
# pca_human, dplyr are available

# create a summary of pca_human
s <- summary(pca_human)

# print out the summary
s

# rounded percentages of variance captured by each PC
pca_pr <- round(100*s$importance[2,], digits = 1) 

# print out the percentages of variance
pca_pr

# create object pc_lab to be used as axis labels
pc_lab <- paste0("PC", 1:8, " (", pca_pr, "%)")

# draw a biplot
biplot(pca_human, cex = c(0.5, 0.8), col = c("grey40", "deeppink2"), xlab = pc_lab[1], ylab = pc_lab[2])

```

*** =sct
```{r}
test_output_contains("s")
test_object("pca_pr", incorrect_msg = "Please adjust the code to change proportions into percentages and then round to onw digit.")
test_output_contains("pca_pr")
test_object("pc_lab")

test_error()
success_msg("Very nice work!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:51411e6d45
## It's tea time!

The [Factominer](https://cran.r-project.org/web/packages/FactoMineR/index.html) package contains a `tea` dataset and it contains the answers of a questionnaire on tea consumption. 

Let's dwell in teas for a bit!

*** =instructions
- Look at first six rows of the `tea` dataset
- Create the `keep_columns` object. Then `select()` the columns from `tea` to create a new dataset. Save the new data as `tea_time`. 
- Look at the summaries and structure of the `tea_time` data.
- visu !

*** =hint
- You can see the first six rows with the `head()` function

*** =pre_exercise_code
```{r}
library(dplyr)
tea <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/tea.csv", sep  =",", header = T)

```

*** =sample_code
```{r}
# the tea dataset and packages FactoMineR and dplyr are available

# look at the head of the tea dataset


# column names to keep in the dataset
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")

# select the 'keep_columns' to create a new dataset
tea_time <- select(tea, "change me!")

# look at the summaries and structure of the data



# visualize the dataset
#joku plot

```

*** =solution
```{r}
# the tea dataset and packages FactoMineR and dplyr are available

# look at the head of the tea dataset
head(tea)

# column names to keep in the dataset
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")

# select the 'keep_columns' to create a new dataset
tea_time <- select(tea, one_of(keep_columns))

# look at the summaries and structure of the data
summary(tea_time)
str(tea_time)

# visualize the dataset


```

*** =sct
```{r}
test_object("tea_time", incorrect_msg = "Did you create `tea_time`?")

test_function("summary", args = "object",incorrect_msg = "Did you look at the summary of `tea_time`?")
test_function("str", args = "object",incorrect_msg = "Did you look at the structure of `tea_time`?")

# test_output_contains("", incorrect_msg = "")
# 
test_error()
success_msg("Nice! Pour yourself a cup of tea and continue to the next exercise :) ")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:42c9b3c01b
## Multiple Correspondence analysis

Exercise info here

*** =instructions
- Do multiple correspondence analysis with the function `MCA()`. Give `tea_time` as the functions first argument. Note that the `MCA()` function visualizes the analysis by default, and the plots can be turned off with the argument `graph = FALSE`.
- Look at the summary of the model.
- Plot the variables of the model. You can either plot the variables or the individuals or both. You can change which one to plot with the `invisible` argument.
- Adjust the code: add argument `habillage = "quali"` (how french!) to the plot. Do you notice what changes?

*** =hint
- See `?MCA` and/or `?plot.MCA`

*** =pre_exercise_code
```{r}
tea_time <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/tea_time.csv", sep  =",", header = T)

sourceDir <- function(path, trace = TRUE, ...) {
    for (nm in list.files(path, pattern = "\\.[RrSsQq]$")) {
       if(trace) cat(nm,":")           
       source(file.path(path, nm), ...)
       if(trace) cat("\n")
    }
}

sourceDir("https://raw.githubusercontent.com/cran/FactoMineR/master/R/")

```

*** =sample_code
```{r}
# tea_time is available

# multiple correspondence analysis
mca <- MCA("change me!", graph = FALSE)

# summary of the model


# visualize MCA
#plot("change me!", invisible=c("ind"))

```

*** =solution
```{r}
# tea_time is available

# multiple correspondence analysis
mca <- MCA(tea_time, graph = FALSE)

# summary of the model
summary(mca)

# visualize MCA
plot(mca, invisible=c("ind"), habillage = "quali")

```

*** =sct
```{r}
test_function("summary", args = "object",incorrect_msg = "Did you look at the summary of the MCA model?")
test_object("mca", incorrect_msg = "Did you create the `mca` object?")
test_error()
success_msg("Dimensions are decreasing fewer by the minute! Good work!")
```
