---
title: 'Analysis of longitudinal data [DRAFT]'
description: 'Repeated measure data'
---

## Meet and Repeat [DRAFT]

```yaml
type: NormalExercise
key: f9ad1f5b4a
lang: r
xp: 100
skills: 1
```

Welcome to the *Analysis of longitudinal data* chapter.

THIS CHAPTER IS A DRAFT

`@instructions`
- Read the `BPRS` data into memory
- Print out the (column) names of the data
- Look at the structure of the data
- Print out summaries of the variables in the data

`@hint`
- Use `str()` to see structure
- Use `summary()` to compute summaries

`@pre_exercise_code`
```{r}

```

`@sample_code`
```{r}
# read the BPRS data
bprs <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)

# look at the (column) names of human
names(bprs)

# look at the structure of human


# print out summaries of the variables


```

`@solution`
```{r}
# read the human data
bprs <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)

# look at the (column) names of human
names(bprs)

# look at the structure of human
str(bprs)

# print out summaries of the variables
summary(bprs)
```

`@sct`
```{r}

test_function("str", args = "object",incorrect_msg = "Please look at the structure of the BPRS data frame")
test_function("summary", args = "object",incorrect_msg = "Please print out summaries of the variables in the BPRS data frame")

test_error()
success_msg("Awsome, you have mastered an important first step of data analysis!")
```

---

## Graphical displays of longitudinal data: The magical gathering

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
key: e9b52fd18b
```

Exercise 2 introduction: Converting to long form

`@instructions`
- Instructions

`@hint`
- Hints

`@pre_exercise_code`
```{r}
library(dplyr)
bprs <- read.table("https://assets.datacamp.com/production/repositories/708/datasets/5346751c337348a8fad84fd239b6b42a730a99c9/BPRS.txt", sep  =" ", header = T)
```

`@sample_code`
```{r}
# dplyr package and bprs are available

```

`@solution`
```{r}
# dplyr package and bprs are available

# Factor treatment & subject
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)

# Convert to long form
BPRSL <- gather(BPRS, key = weeks, value = bprs, -treatment, -subject)
BPRSL <-  mutate(BPRSL, week = as.integer(substr(weeks,5,5)))

# Let's see the gathered "long" data
glimpse(BPRSL)
```

`@sct`
```{r}

test_function("gather", args = "object",incorrect_msg = "Did you use the gather function?")
test_function("mutate", args = "object",incorrect_msg = "Did you use the mutate function")

test_error()
success_msg("Success message!")
```
