---
title: "Credic card data (cc) exploration"
author: "Tuomo Nieminen"
date: "19 joulukuuta 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Explore and analyze the credit card data.

### Read  the data

```{r}
setwd("C:/Users/Tuomo/Documents/GitHub/Helsinki-Open-Data-Science/data")
cc <- read.csv("cc.csv")
```


# Exploration

Dimensions of the data, basic stats of variables, exploratory graphs.

## 1. Structure

```{r}

# dimensions (N x d)
dim(cc)

# structure
str(cc)

```

## 2. Summaries

```{r}
# summaries of transaction amounts and frauds
summary(cc$Amount)
```


Class == 1 denotes a fraudulent transaction.

```{r}
table(cc$Class)
```

```{r}
# how many zero amount transactions?
sum(cc$Amount == 0)

# exclude rows where Amount equals zero
cc <- cc[cc$Amount != 0, ]
```

## 3. Exploring Amount vs frauds (1)

```{r}
# recode 'Class' as a factor and save as a column called 'fraud'
cc$fraud <- factor(cc$Class, levels = c(0,1), labels = c("N","Y"))

# box plot of transaction amounts by fraud/legit on log scale
library(ggplot2)
p <- ggplot(cc, aes(x = fraud, y = Amount))
box <- p + geom_boxplot()
box + scale_y_log10()
```

## 4. Exploring Amount vs frauds (2)

```{r}
# categorize transaction amount
cc$amount <- cut(cc$Amount, breaks = c(0, 5, 20, 100, 1000, 5000))

# create a function which computes a percentage table
prc_tbl <- function(x, y) {
  t1 <- table(x, y)
  p <- prop.table(t1)
  p <- addmargins(p)
  p <- round(p, 4)
  100*p
}

# percentage table of amount categories vs fraud/non-fraud
prc_tbl(cc$amount, cc$fraud)
```

# Logistic regression modelling

Fitting a logistic regression model with glm(). Interpreting the parameters. Calculating prediction accuracy. k-fold cross validation.

## 5. fitting a logistic regression model

```{r}
# fit a model
m <- glm(fraud~Amount+V2+V3+V4, family = binomial(), data = cc)

# summary of the model
summary(m)
```


## 6. analyzing the model output: coefficients

```{r}
# print the coefficients
m

# print exponentiated coefficients
exp(coefficients(m))
```


## 7. model performance: prediction accuracy

```{r}
# predict the response
cc$prediction <- predict(m, type = "response")

head(cc[,c("Class","prediction")])

# average predicted fraud probabilities for frauds and nonfrauds
by(cc$prediction, cc$fraud, mean)
```

## 8. model performance: prediction accuracy (2)
```{r}
# prediction accuracy by simply guessing vs our model 

# cost function (how many times we predict the wrong class)
cost <- function(prediction, class) sum(abs(prediction - class) > 0.5)

# guess: predict everything as non-fraud
incorrect_guesses <- cost(prediction = 0, class = cc$Class)
incorrect_guesses

accuracy_guess <- 1 - incorrect_guesses / nrow(cc)
accuracy_guess * 100

# how did the model perform?
incorrect_predictions <- cost(prediction = cc$prediction, class = cc$Class)
accuracy_model <- 1 - incorrect_predictions / nrow(cc)
accuracy_model * 100
```


## 9. model performance: cross-validation

```{r}
library(boot)

# cost function
cost <- function(prediction, class) sum(abs(prediction - class) > 0.5)

# k-fold cross-validation
cv <- cv.glm(data = cc, cost = cost, glmfit = m, K = 10)$delta

accuracy <- 1 - cv[1] / nrow(cc)
accuracy*100

```

